\chapter*{Preface}
\addcontentsline{toc}{chapter}{Preface}

This MSc thesis has been carried out within the Embedded and Networked Systems group at the Delft University of Technology under the supervision of Dr. Przemysław Pawełczak and Ph.D. Candidate Jasper de Winkel. The goal from the beginning was always to improve upon the state-of-the-art battery-less system called FreeBie, developed by De Winkel and Dr. Pawełczak. While familiarizing myself with FreeBie and Bluetooth Low-Energy (BLE), it slowly became evident why the current revision of BLE was not perfectly suited for intermittent devices. This gave me motivation and a clear vision for improving BLE for use within the intermittent domain. 

\vspace{1\baselineskip}

In the first year of my MSc, I followed a course called ``Wireless IoT and Local Area Networks'' by Dr. Pawełczak, which quickly became one of my favorite courses I got to experience at TU Delft. During this course, Dr. Pawełczak presented his research on intermittent devices, which led me to do my thesis with him, with Jasper de Winkel becoming my daily supervisor. Firstly, I would like to thank them both for this opportunity. I would also like to thank them for their continued effort, meticulous eye for detail, and expertise. Finally, and perhaps more importantly, I would like to thank them for what has been a very enjoyable and personal experience during the final year of my MSc journey. To Dr. Asterios Katsifodimos, thank you for accepting the role of committee member and reviewing my work as part of it; I sincerely hope you enjoy it.

Looking back at my time at TU Delft and my BSc before, I am filled with pride and the simultaneous feeling that it has been so long, and it is already coming to an end.

\vspace{1\baselineskip}

\reportAuthor

\vspace{1\baselineskip}

\noindent
Delft, The Netherlands

\noindent
\today